Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-17 paper/paper.tex
perfect-partition

In Section~\ref{sec:ppl}, we study a restricted model of balanced re-partition, called perfect partition.
For that, we present a $O(k\cdot \ell)$-competitive algorithm, which is optimal.
We believe that this result is an important building block on the way to an optimal algorithm for the general model.

@@

\section{An Optimal Algorithm for Perfect Partition}	\label{sec:ppl}
%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
%\maciek{ every algorithm must colocate every communicating pair.
%	\mahmoud{updated!}}
%\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}

@@

A \emph{perfect partition} is a partitioning of nodes into $\ell$ clusters
such that it incurs no inter-cluster communication throughout the input sequence.
Equivalently, a perfect partitioning it an $\ell$-way cut of the communication graph of the input sequence.
In this section,
we give an algorithm for a restricted variant of  \OBRP{},
%\maciek{This is not sufficient! We also disallow serving requests remotely.}
where an optimal offline algorithm ($\OPT$) moves to this perfect partition
at the beginning and stays there perpetually.
%\maciek{Why do we have to assume that? This is just what OPT does, i.e., we can prove that this strategy (a) lower bounds the OPT's cost and (b) is sufficent for our analysis. But now is sounds like we restrict the model even further.}
We refer this variant as \PPOBRP{}.
The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
%
%\maciek{This is not a valid description of the model. We need the assumption that collocation is needed in our model. Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
The first request between any two non-collocated nodes reveals an edge of the communication graph.
Our online algorithm collocates them immediately and never separates them since by assumption \OPT does the same.
Revealed edges, up to any point in the sequence,
induce some connected subgraphs and we refer to them as \emph{components}.
An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component-respecting}.

\subsection{Perfect Partition Learner (\PPL) Algorithm}

@@

\begin{algorithm}
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \begin{algorithmic}
%        \Require 
%        $k, \ell$,
%        initial configuration $P_I$,
%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
%        \Ensure A final configuration $P_F$ 
        \STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
        \STATE{$P_0 := P_I$}
         \label{line:initcomponents}
        \FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
        \STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
        \IF{$C_1 \neq C_2$}
        \STATE {Unite the two components into a single component $C'$ and
        $\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
        \IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
	     \COMMENT{i.e.~if not in the same cluster}    
    	}       
        \STATE {$P_{t} = \mathit{re\mhyphen partition}(P_{t-1}, P_I, \mathcal{C})$} 
         \COMMENT{move to $P$ closest to $P_I$}
        \label{line:rebalance} 
        \ENDIF
        \ENDIF
        \ENDFOR
    \end{algorithmic}
    \caption{Perfect Partition Learner (\PPL)}
    \label{alg:ppl}
      \end{algorithm}

%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect partition closest to $P_I$.
Hence it never moves to a configuration beyond distance $\Delta$.      
\begin{property} \label{prop:dist<OPT}
    Let $P$ be any configuration chosen by \PPL at Line $\ref{line:rebalance}$.
    Then, $\dist(P,P_I) \leq \Delta$.
\end{property}

@@

 In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
 Note that the lower bound holds also in the perfect partition model, as the constructed input sequence allows \OPT to move to a perfect partition.
 The corollary is that PPL is optimal.

>>>>

In Section~\ref{sec:ppl}, we study a restricted model of balanced re-partition, called perfect-partition.
For that, we present a $O(k\cdot \ell)$-competitive algorithm, which is optimal.
We believe that this result is an important building block on the way to an optimal algorithm for the general model.

@@

\section{An Optimal Algorithm for Perfect-Partition}	\label{sec:ppl}
%\subsection{Perfect Partition Model}\label{sec:perfectPartition}
%\maciek{Here introduce a restricted model of \PPOBRP{}, the restricted variant where the input sequence can be perfectly partitioned}
%\maciek{ every algorithm must colocate every communicating pair.
%	\mahmoud{updated!}}
%\maciek{Note that the algorithm from Section~\ref{sec:k3} is $O(\ell)$-competitive for the Perfect Partition Model, as during the first phase, a perfect partition of input exists.}

@@

A \emph{perfect-partition} is a partitioning of nodes into $\ell$ clusters
such that it incurs no inter-cluster communication throughout the input sequence.
Equivalently, a perfect-partitioning it an $\ell$-way cut of the communication graph of the input sequence.
In this section,
we give an algorithm for a restricted variant of  \OBRP{},
%\maciek{This is not sufficient! We also disallow serving requests remotely.}
where an optimal offline algorithm ($\OPT$) moves to this perfect-partition
at the beginning and stays there perpetually.
%\maciek{Why do we have to assume that? This is just what OPT does, i.e., we can prove that this strategy (a) lower bounds the OPT's cost and (b) is sufficent for our analysis. But now is sounds like we restrict the model even further.}
We refer this variant as \PPOBRP{}.
The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect-partition while not paying too much relative to \OPT.
%
%\maciek{This is not a valid description of the model. We need the assumption that collocation is needed in our model. Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
The first request between any two non-collocated nodes reveals an edge of the communication graph.
Our online algorithm collocates them immediately and never separates them since by assumption \OPT does the same.
Revealed edges, up to any point in the sequence,
induce some connected subgraphs and we refer to them as \emph{components}.
An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component-respecting}.

\subsection{Perfect-Partition Learner (\PPL) Algorithm}

@@

\begin{algorithm}
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \begin{algorithmic}
%        \Require 
%        $k, \ell$,
%        initial configuration $P_I$,
%        sequence of  requests $\sigma_1, \dots, \sigma_N$ 
%        \Ensure A final configuration $P_F$ 
        \STATE {For each node $v$ create a singleton component $C_v$ and add it to $\mathcal{C}$}
        \STATE{$P_0 := P_I$}
         \label{line:initcomponents}
        \FOR {each  request $\sigma_t=\{u,v\}, 1 \leq t \leq N$}
        \STATE Let $C_1 \ni u$ and $C_2 \ni v$ be the container components
        \IF{$C_1 \neq C_2$}
        \STATE {Unite the two components into a single component $C'$ and
        $\mathcal{C} = (\mathcal{C}\setminus\set{C_1, C_2}) \cup ~\set{C'}$} \label{line:mergecomponents}
        \IF{$\mathit{cluster}(C_1, P_{t-1}) \neq \mathit{cluster}(C_2, P_{t-1})$
	     \COMMENT{i.e.~if not in the same cluster}    
    	}       
        \STATE {$P_{t} = \mathit{re\mhyphen partition}(P_{t-1}, P_I, \mathcal{C})$} 
         \COMMENT{move to $P$ closest to $P_I$}
        \label{line:rebalance} 
        \ENDIF
        \ENDIF
        \ENDFOR
    \end{algorithmic}
    \caption{Perfect-Partition Learner (\PPL)}
    \label{alg:ppl}
      \end{algorithm}

%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
      We note that the $\mathit{re\mhyphen partition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a perfect-partition closest to $P_I$.
Hence it never moves to a configuration beyond distance $\Delta$.      
\begin{property} \label{prop:dist<OPT}
    Let $P$ be any configuration chosen by \PPL at Line $\ref{line:rebalance}$.
    Then, $\dist(P,P_I) \leq \Delta$.
\end{property}

@@

 In Section~\ref{sec:lowerbound} we constructed a $\Omega(k \cdot \ell)$ for \OBRP{}.
 Note that the lower bound holds also in the perfect-partition model, as the constructed input sequence allows \OPT to move to a perfect-partition.
 The corollary is that PPL is optimal.
